---
title: "Homework 2 Report"
output: html_document
---


## Name

Bhavik Singhal

### Setup

Read in the data with `read_csv()` and store the data as an R object named `dataset`. Check the data to make sure all of the expected observations and variables are there.


```{r}
## Load the data and any necessary packages here.
library(dplyr)
library(tidyverse)

dataset = read_csv("maacs.csv.gz")
```

## Part 1

We will first consider the relationship between FEV1 and age. In general, it is expected that as children get older (and hence, larger in size), their FEV1 values should get higher. 

Consider the statement "FEV1 values in children are higher in older children relative to younger children". 

Write a function in R that takes the `dataset` object as an argument and returns `TRUE` if the statement above is true for the dataset and `FALSE` otherwise. 

NOTE: In order to write this function, you will need to translate the statement above into something that can be checked with the data. There are many ways in which you can do that translation correctly and you only need to pick one way here. 

NOTE: For this part, do not use any plots.

```{r}
## Write your function here
correlation_premise <- function(dataset){
  test <- cor.test(dataset$age, dataset$fev1, method = "pearson", alternative = "greater")
  
  (test$estimate > 0) && (test$p.value < 0.05)
}

correlation_premise(dataset)
```


## Part 2

Fit a linear regression model with FEV1 as the outcome and age as a predictor.

How much does FEV1 change for a 1-year increase in the child's age? 

```{r}
## Add your code here
lin_fit = lm(fev1 ~ age, data = dataset)
summary(lin_fit)
coef(lin_fit)["age"]
```


Write your data analysis statement interpreting the regression model here: 

With a 1 year increase in the child's age, there is a 0.17 liter increase in FEV on average, according to the linear regression fit.

## Part 3

Develop **three** supporting premises derived from the data that support the statement you wrote in Part 2. These can be plots, other summary statistics, or model results. 

NOTE: 

* At least one supporting premise should use a plot.

* Do not use the code you write in Part 1 as a supporting premise

```{r}
## Add your code here

```


Write the three supporting premise statements here:

1. A one-sample two-sided t-test shows that older children have a significantly higher mean fev1 than younger children.

2. Comparing the boxplots of fev1 for both older and younger children, a higher median is seen in the older children boxplot compared to the younger children boxplot.

3. A histogram with older children fev1 values for older children shows a distribution centered at a higher value while the histogram for the younger children will be centered at a lower value. Histograms are also both bell-shaped.


## Part 4

For each of the supporting premises above, write a function that takes the `dataset` object as an argument and returns `TRUE` if the supporting premise statement above is true for the dataset and `FALSE` otherwise. 

For statements involving plots, instead of returning `TRUE` or `FALSE`, you function should do two things:

1. Produce the plot that is used in the statement

2. Produce a hypothetical version of the plot in the event that the statement is true. This can be done using simulated data or by simply hand drawing a plot.


```{r}
## Function for supporting premise statement 1 (T-Tests)

ttest_premise = function(dataset){
  median_age = median(dataset$age, na.rm = TRUE)
  young = dataset$fev1[dataset$age <= median_age]
  old = dataset$fev1[dataset$age >  median_age]
  
  test_results = t.test(old, young, alternative = "greater")
  
  (mean(old, na.rm = TRUE) > mean(young, na.rm = TRUE))
}
```


```{r}
## Function for supporting premise statement 2 (boxplot)
boxplot_premise = function(dataset){
  
  dataset$AgeCategory = ifelse(dataset$age > median(dataset$age, na.rm = TRUE), "Older", "Younger")

  #Normal
  
  boxplot(fev1 ~ AgeCategory, data = dataset, main = "Distribution of FEV1 by Age", col = c("red", "green"))
  
  #Hypothetical
  
    simulation = data.frame( AgeCategory = rep(c("Younger","Older"), each = 50), fev1 = c(rnorm(50, mean = 1.5, sd = 0.3), rnorm(50, mean = 3, sd = 0.3)))
    
  boxplot(fev1 ~ AgeCategory, data = simulation, main = "Distribution of FEV1 Hypothetical by Age", col = c("red", "green"))
}
```


```{r}
## Function for supporting premise statement 3
histogram_premise = function(dataset){
  
  median_age = median(dataset$age, na.rm = TRUE)
  dataset$AgeCategory = ifelse(dataset$age > median_age, "Older", "Younger")
  
  #Normal
  
  young = dataset$fev1[dataset$AgeCategory == "Younger"]
  old = dataset$fev1[dataset$AgeCategory == "Older"]
  
  hist(young, main = "Younger Observed Scenario", col = "red")
  hist(old, main = "Younger Observed Scenario", col = "green")
  
  #Hypothetical
  
  sim_young = rnorm(50, mean = 1.5, sd = 0.3)
  sim_old   = rnorm(50, mean = 3, sd = 0.3)
  
  hist(sim_young, main = "Younger Hypothetical Scenario", col = "red")
  hist(sim_old, main = "Older Hypothetical Scenario", col = "green")
}
```

Execute each of your function and show that the produce the expected output.

```{r}
ttest_premise(dataset)
boxplot_premise(dataset)
histogram_premise(dataset)
```

## Part 5

Describe one alternative to the primary statement "FEV1 values in children are higher in older children relative to younger children". 
FEV1 values in children are lower in older children relative to younger children.

Create a fault tree for the alternative outcome describing how the alternative outcome could be realized in the data even if the primary statement were true.

Your fault tree should be created as a separate image and does not need to be created in R. Upload the image of the fault tree to Canvas.


